---
title: "DDS Final Project - Attrition"
author: "Josh Turk"
date: "2023-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries Used
```{r, warning=FALSE}
library(tidyverse)
library(caret)
```

Data Import and Cleaing
```{r}
data <- read.csv("Datasets/CaseStudy2-data.csv", header = TRUE)

#Remove columns that aren't useful
data <- subset(data, select = -c(ID,EmployeeCount,EmployeeNumber,Over18,StandardHours))

#Change all character variables to factors
data[sapply(data, is.character)] <- lapply(data[sapply(data, is.character)], 
                                       as.factor)

#Create binary var for attrition
data$AttritionCoded <- with(data, ifelse(Attrition == "Yes", 1, 0))

```

Creating Train and Test Sets
```{r}
split <- 0.75
trainIndex <- sample(seq(1,870,1),870*split)
dataTrain <- data[trainIndex,]
dataTest <- data[-trainIndex,]
```

Here we are building a kNN model to predict attrition. We will use 10-fold cross validation and upsample the data to create a balanced dataset. 
```{r}

ctrl <- trainControl(method = "repeatedcv", 
                     number = 10, 
                     repeats = 5, 
                     verboseIter = FALSE,
                     sampling = "up")

model <- train(Attrition ~ Age+
                 EducationField+EnvironmentSatisfaction+
                 Gender+HourlyRate+JobLevel+
                 JobRole+JobSatisfaction+JobInvolvement+
                 MonthlyIncome+MonthlyRate+NumCompaniesWorked+
                 OverTime+RelationshipSatisfaction+WorkLifeBalance+
                 YearsSinceLastPromotion+YearsWithCurrManager,
               data = dataTrain,
               method = "knn",
               trControl = ctrl,
               preProcess = c("center","scale"))

predict <- predict(model,newdata = dataTest)
confusionMatrix(table(predict,dataTest$Attrition))
```

Attrition Rate by Job Role
```{r}
#create dataset
jobdat <- data %>% group_by(JobRole) %>% 
  summarize(percAttr = sum(AttritionCoded)/length(AttritionCoded)*100)

#plot data
jobdat %>% ggplot(aes(reorder(JobRole, percAttr), percAttr)) + 
  geom_col(fill = "#303c94") + coord_flip() + labs(y = "% Attrition",  x= "Job Role")
```

Monthly Income Regression
```{r}
fit3 <- lm(MonthlyIncome ~ JobLevel+TotalWorkingYears+
             JobRole+JobRole*JobLevel, data = data)
summary(fit3)
```

Job Satisfaction by Attrition T-Test
```{r}
#job satisfaction by attrition, ttest
x <- data %>% filter(Attrition == "Yes")
x <- x[,"JobSatisfaction"]
y <- data %>% filter(Attrition == "No")
y <- y[,"JobSatisfaction"]

t.test(x,y)
```

Monthly Income by Attrition T-Test
```{r}
#montlyincome by attrition, ttest
a <- data %>% filter(Attrition == "Yes")
a <- a[,"MonthlyIncome"]
b <- data %>% filter(Attrition == "No")
b <- b[,"MonthlyIncome"]

t.test(a,b)
```

YouTube Presentation: https://youtu.be/b8qVZO8sTnY
RShiny App:  https://josh-turk.shinyapps.io/AttritionProject/

